# Examples

For regression, let's use the Tecator data in the modeldata package:

```{r, message=FALSE}
library(tidymodels)
library(plsmod)
tidymodels_prefer()
theme_set(theme_bw())

data(meats, package = "modeldata")
```

Note that using `tidymodels_prefer()` will resulting getting `parsnip::pls()` instead of `mixOmics::pls()` when simply running `pls()`.

Although plsmod can fit multivariate models, we'll concentration on a univariate model that predicts the percentage of protein in the samples.  

```{r}
meats <- meats |> select(-water, -fat)
```

We define a sparse PLS model by setting the `predictor_prop` argument to a value less than one. This allows the model fitting process to set certain loadings to zero via regularization. 

```{r}
sparse_pls_spec <- 
  pls(num_comp = 10, predictor_prop = 1/3) |> 
  set_engine("mixOmics") |> 
  set_mode("regression")
```

The model is fit either with a formula or by passing the predictors and outcomes separately:

```{r}
form_fit <- 
  sparse_pls_spec |> 
  fit(protein ~ ., data = meats)
form_fit

# or 

sparse_pls_spec |> 
  fit_xy(x = meats |> select(-protein), y = meats$protein)
```
The `pls()` function can also be used with categorical outcomes. 
