---
title: "plsmod"
---

The plsmod package serves as an extension to the parsnip package that enables tiydmodels users to fit several types of [partial least squares models](https://scholar.google.com/scholar?hl=en&as_sdt=0%2C7&q=%22partial+least+squares+models%22&btnG=) (PLS). If you are unfamiliar with tidymodels, please see the [beginning help pages as tidymodels.org](https://www.tidymodels.org/start/). 

With tidymodels, there are possible _engines_ that can be used to fit a particular model. For PLS, the possible models are: 

```{r}
library(plsmod)
show_engines("pls")
```

## An example

For a demonstration of regression modeling, well use the [Tecator data](https://modeldata.tidymodels.org/reference/meats.html) in the modeldata package:

```{r, message=FALSE}
library(dplyr)
data(meats, package = "modeldata")
```

Note that using `tidymodels_prefer()` will resulting getting `parsnip::pls()` instead of `mixOmics::pls()` when simply running `pls()`.

Although plsmod can fit multivariate models, we'll concentration on a univariate model that predicts the percentage of protein in the samples.  

```{r}
meats <- meats %>% select(-water, -fat)
```

We define a sparse PLS model by setting the `predictor_prop` argument to a value less than one. This allows the model fitting process to set certain loadings to zero via regularization. 

```{r}
sparse_pls_spec <- 
  pls(num_comp = 10, predictor_prop = 1 / 3) %>% 
  set_engine("mixOmics") %>% 
  set_mode("regression")
```

The model is fit either with a formula or by passing the predictors and outcomes separately:

```{r}
form_fit <- 
  sparse_pls_spec %>% 
  fit(protein ~ ., data = meats)
form_fit

# or 

sparse_pls_spec %>% 
  fit_xy(x = meats %>% select(-protein), y = meats$protein)
```

The `pls()` function can also be used with categorical outcomes, provided that the outcome column is an R factor vector. 

The number of components and the amount of sparsity can be optimized using the tune package. See Chapter 17 of [_Tidy Models with R_](https://www.tmwr.org/tuning.html) for more information and examples on how to tune model hyperparameters using tidymodels. 

For prediction, the basic `predict()` method can be used: 

```{r}
predict(form_fit, head(meats))
```
